---
title: "GSA Exploration"
author: "Ted Wroblewski"
date: "Thursday, April 14, 2016"
output: html.document
---

Blatantly stole this to load in all of the FPDS Full Files:

```{r}
require(dplyr)
require(ggplot2)
require(pander)
require(knitr)
opts.chunk$set(tidy=TRUE)


# Load data into single frame from multiple CSV files in directory.
# Directory of files from Kristin, filenames contain numerical date
# information that we will want to capture.
datadir <- "C:/Users/517951/OneDrive - Booz Allen Hamilton/GSA Competition"
reading.helper <- function(infile, datadir, ...) {
## Function to read in csv and strip date information from file name
temp <- read.csv(paste(datadir, infile, sep = "/"), stringsAsFactors = F,
...)
temp$period <- gsub("FPDSFULL", "", infile) %>% gsub("\\.csv", "", .)
return(temp)
}
# if (exists(fpds)){rm(fpds)}
files <- list.files(datadir, pattern = "FPDSFULL[0-9.]*.csv")
fpds <- do.call(rbind, lapply(files, function(x) reading.helper(x, datadir)))
# Here are the column headers in the data.frame
names(fpds)
```

Whittled down to keep only the variables I would like

```{r}
fpds.vars <- c("Contracting.Agency.ID", "Contracting.Agency.Name",
               "Contracting.Department.ID","Contracting.Department.Name",
               "Contracting.Office.ID", "Contracting.Office.Name", 
               "Modification.Number", "PIID","PIID.Agency.ID",
               "Referenced..IDV.PIID", "Referenced.IDV.Agency.ID",
               "Referenced.IDV.Mod.Number","Solicitation.ID","Completion.Date",
               "Date.Signed", "Effective.Date","Est..Ultimate.Completion.Date",
               "Fiscal.Year", "Funding.Agency.ID","Funding.Agency.Name",
               "Funding.Department.ID","Funding.Department.Name","Funding.Office.ID",
               "Funding.Office.Name","Type.of.Contract","NAICS.Code",
               "NAICS.Description", "Product.or.Service.Code",
               "Product.or.Service.Description", 
               "Principal.Place.of.Performance.City.Name", 
               "Principal.Place.of.Performance.Country.Code",
               "Principal.Place.of.Performance.Country.Name",
               "Principal.Place.of.Performance.County.Code",
               "Principal.Place.of.Performance.County.Name",
               "Principal.Place.of.Performance.Location.Code",
               "Principal.Place.of.Performance.State.Code",
               "Extent.Competed", "Number.of.Offers.Received","Type.of.Set.Aside",
               "Award.or.IDV.Type", "Contractor.Name", "Doing.Business.As.Name",
               "DUNS.Number","Global.DUNS.Number", "Global.Vendor.Name",
               "IDV.Who.Can.Use", "Base.and.Exercised.Options.Value",
               "Action.Obligation", "Base.and.All.Options.Value", "period")

fpds.mine <- fpds[fpds.vars]

```

Creating a PoP variable -- NEED TO FIGURE OUT WHY NEGATIVE POPS
```{r}
fpds.mine$End.Date <- as.Date(fpds.mine$Est..Ultimate.Completion.Date,"%m/%d/%Y")
fpds.mine$Start.Date <- as.Date(fpds.mine$Effective.Date,"%m/%d/%Y")

fpds.mine$pop <- difftime(fpds.mine$End.Date,fpds.mine$Start.Date, units="weeks")
```

just getting a feel for how many unique instances there are of certain variables

```{r}
attach(fpds.mine)

unique(Contracting.Agency.Name)
unique(Contracting.Office.Name)
unique(Funding.Agency.Name)
```

Now, I want to line up everything to the contract/DO level and see mods,
value (NEED TO EXPERIMENT BETWEEN CEILING/ALL OPTIONS, FUNDING/EXERCISED)

```{r}
library(sqldf)
library(lucr)

#Creating a unique ID for each contract

fpds_mine <- mutate(fpds.mine, uniqueId = paste(PIID.Agency.ID, PIID, 
                                                Referenced.IDV.Agency.ID,
                                                Referenced..IDV.PIID, sep='-'))

colnames(fpds_mine)[colnames(fpds_mine)=="Product.or.Service.Code"] <- 
  "Product.Service.Code"

colnames(fpds_mine)[colnames(fpds_mine)=="Product.or.Service.Description"] <- 
  "Product.Service.Description"

names(fpds_mine) <- gsub("\\.", "", names(fpds_mine))


#Should also look at adding in further down to offices

fpds_contract <- sqldf("SELECT uniqueID, ContractingAgencyID,
                       ContractingAgencyName, 
                       FundingAgencyID,FundingAgencyName, TypeofContract, 
                       NAICSCode,NAICSDescription, ProductServiceCode,
                       ProductServiceDescription, TypeofSetAside,
                       GlobalDUNSNumber, GlobalVendorName,
                       PrincipalPlaceofPerformanceStateCode,
                       PrincipalPlaceofPerformanceCountyCode,
                       Count(ModificationNumber) as Mods,
                       max(BaseandAllOptionsValue) as Value, max(pop) as PoP,
                       max(NumberofOffersReceived) as Offers 
                       FROM fpds_mine 
                       GROUP BY uniqueID, ContractingAgencyID, 
                       ContractingAgencyName, FundingAgencyID, 
                       FundingAgencyName, TypeofContract, NAICSCode,
                       NAICSDescription, ProductServiceCode,
                       ProductServiceDescription, TypeofSetAside, 
                       GlobalDUNSNumber, GlobalVendorName,
                       PrincipalPlaceofPerformanceStateCode,
                       PrincipalPlaceofPerformanceCountyCode", stringsAsFactors=TRUE)

fpds_contract$Value2 <- from_currency(as.vector(fpds_contract$Value), decimal_delim = ".")

fpds_contract$Value3 <- ifelse(grepl("-", fpds_contract$Value,), 
                              -1*fpds_contract$Value2, fpds_contract$Value2)
#Need to drop Value and Value2 fields in analysis                       
```

I did a lot of data exploration in rattle and tried to see if I could work a 
clustering algorithm to no avail.

Export to csv

```{r}

write.csv(fpds_contract, 
          "C:/Users/517951/OneDrive - Booz Allen Hamilton/GSA Competition/fpds.csv")
